---
title: "GartleyHW6"
author: "Roy Gartley"
date: "2026-04-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyr)
library(dplyr)
```



```{r }
load("house_prices.rda")
house_prices <- house_prices %>%
  mutate(year_num = as.numeric(format(as.Date(date), "%Y")))
# --- (a) Line plots for house_price_index ---
plot_a <- ggplot(house_prices, aes(x = year_num, y = house_price_index)) +
  geom_line() +
  facet_wrap(~state) +
  scale_x_continuous(
    breaks = c(1980, 2000, 2020),
    labels = c("'80", "'00", "'20"),
    # without limits 20 does not show 
    limits = c(1975, 2022) 
  ) +labs(title = "House Price Index Trend", x = "Year", y = "HPI")+
  theme_minimal() +
  #needed to modify the text size to get the years to show 
  theme(axis.text.x = element_text(size = 7))
print(plot_a)
```

## Comment for this set

Having 50 plots is a lot to look at, but it does clearly show that there are uniqe relationships between house prices and the year for each state.
With unique responses around the 2008 houseing crisis some subtle and some not so much.

```{r}
house_reshaped <- house_prices %>%
  gather(key = "measure", value = "value", -year_num, -state, -date)
# --- (c) Replicate graph with two lines ---
plot_c <- ggplot(house_reshaped, aes(x = year_num, y = value, color = measure)) +
  geom_line() +
  facet_wrap(~state) +
  scale_x_continuous(
    breaks = c(1980, 2000, 2020),
    labels = c("'80", "'00", "'20"),
    limits = c(1975, 2022)
  ) +labs(title = "House Price Index Trend", x = "Year", y = "HPI")+
  theme_minimal() +
  #needed to modify the text size to get the years to show 
  theme(axis.text.x = element_text(size = 7))

print(plot_c)
```

## Comment for this set
You cannot pull anything meaningful from these graphs when it comes to correlation.
Unemployment is just too small of a value vs the 100+ value from house price index.
No its not a good set of charts. 
